Reproducible research in archaeology using R & rrtools

### 2018-03-07 analysis/paper/paper.Rmd
Update paper.Rmd
adding a few sentences
- is what I added to "who is 'we'" actually correct?
- if there is no trend in an increase of using R to be seen, how plausible is it, that there is a paradigm shift if a few people start sharing code? I'm just unsure whether we can be this definite with these small numbers...  Though I'd like it, if there was one!

In this presentation we will talk about how archaeologists are changing their data analysis practices to make their research more reproducible and open. We will the rrtools package, which is our specific contribution to help with the challenge of making reproducible research in archaeology easier.

@@

This is an important change away from using mouse-driven, point-and-click software for data analysis. When archaeologists use software such as SPSS and Past (as you see in our screenshots here), it is very difficult for them to share the details of their analysis. When readers cannot see the details of an analysis, it is difficult for them to be confident in the results described in an article.

@@

First, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences. We also see increases in multi-disciplinary journals such as PeerJ and PLOS One.

@@

The fourth detail we see is the distinctive types of journals that have articles citing R. There is a clear focus on scientific archaeology, with _Journal of Archaeological Science_ having the highest total number of articles. We also see geoarchaeology, lithics, and environmental archaeology represented in the journal names. We can investigate this further with statistical analysis of the words in the titles of all the articles in our sample.

>>>>

In this presentation we will talk about how archaeologists are changing their data analysis practices to make their research more reproducible and open. We will introduce the rrtools package, which is our specific contribution to help with the challenge of making reproducible research in archaeology easier. "Our" meaning in this case a group of people whose names you can see here and / or who took part in thethe 2017 Summer School on Reproducible Research in Landscape Archaeology at the Freie Universität Berlin (17-21 July), funded and jointly organized by Excellence Cluster 264 Topoi, the Collaborative Research Center 1266, and ISAAKiel under the guidance of Ben Marwick.

@@

This is an important change away from using mouse-driven, point-and-click software for data analysis. When archaeologists use software such as SPSS and Past (as you see in our screenshots here), it is very difficult for them to share the details of their analysis. When readers cannot see the details of an analysis, it is difficult for them to be confident in the results described in an article. Also, for other researchers it may be near impossible to implement the described analysis in their own work as they lack the details on how to come to the desired conclusions.

@@

First, the areas of science that have seen big growth in the use of R are ecology, evolution and related biological sciences. This is especially significant for archaeologists, as ecology and biology are subjects from which they quite often adapt methods for analysis, as for example in point-pattern-analysis and statistical modelling. 
We also see increases in multi-disciplinary journals such as PeerJ and PLOS One.

@@

The fourth detail we see is the distinctive types of journals that have articles citing R. There is a clear focus on scientific archaeology, with _Journal of Archaeological Science_ having the highest total number of articles. Archaeology as a subject has always been situated between a humanistic and a scientific tradition, but studies using statistical methods lean towards the sciences, which we mean with the term "scientific archaeology". We also see geoarchaeology, lithics, and environmental archaeology represented in the journal names. We can investigate this further with statistical analysis of the words in the titles of all the articles in our sample.
